---
title: "Qiime2_to_Phyloseq"
author: "Cody Glickman"
date: "10/26/2018"
output: html_document
---

## Purpose
This script details the steps to convert [qiime2](https://docs.qiime2.org/2018.8/concepts/) objects into a [Phyloseq](https://joey711.github.io/phyloseq/) object. The workflow of processing data with Qiime2 can be found at the [Moving Pictures](https://docs.qiime2.org/2018.8/tutorials/moving-pictures/) tutorial. This script is adapted from Pedro J. Torres.

## Requirements
Qiime2 artifacts needed to convert for phyloseq analysis:

  * Metadata/Mapping File
  * Table File (.qza)
  * Taxonomy File (.qza)
  * Tree File (.qza) {Optional}
  
## Convert Qiime2 qza on Command Line
The chunk below contains commands executed on the command line. Qiime should be installed and in your path. Otherwise the commands can be copied and pasted into the terminal running a qiime environment.  

```{r}
system('mkdir -p Phyloseq')
#  mkdir -p Phyloseq

# Export table
system("qiime tools export 'Demo_Data/table.qza' --output-dir 'Phyloseq/'")
#  qiime tools export 'Demo_Data/table.qza' --output-dir 'Phyloseq/'

# Export taxonomy
system("qiime tools export Demo_Data/taxonomy.qza --output-dir Phyloseq/")
#  qiime tools export Demo_Data/taxonomy.qza --output-dir Phyloseq/

# Export tree
system("qiime tools export Demo_Data/unrooted-tree.qza --output-dir Phyloseq/")
#  qiime tools export Demo_Data/unrooted-tree.qza --output-dir Phyloseq/
```

The header of the taxonomy file must be converted to match proper format for merging with biom file
```{r}
system("sed 's/Feature ID/#OTUID/' Phyloseq/taxonomy.tsv | sed 's/Taxon/taxonomy/' | sed 's/Confidence/confidence/' > Phyloseq/biom-taxonomy.tsv")

## Remove redundant file
system("rm Phyloseq/taxonomy.tsv")
```

Merge taxonomy with biom file. Qiime in path is also required for this command. The command can be copied and pasted into a terminal with a qiime environment
```{r}
system("biom add-metadata \
    -i Phyloseq/feature-table.biom \
    -o Phyloseq/taxa_table.biom \
    --observation-metadata-fp Phyloseq/biom-taxonomy.tsv \
    --sc-separated taxonomy")

#  biom add-metadata -i Phyloseq/feature-table.biom -o Phyloseq/taxa_table.biom --observation-metadata-fp Phyloseq/biom-taxonomy.tsv --sc-separated taxonomy
```


## Load in Qiime Data
Now the data is ready to use in Phyloseq! Below we load in the biom table, tree file (optional), and mapping file. Then we combine them into a phyloseq object. There is a warning that can be ignored upon importing the biom table. [Tutorial](http://joey711.github.io/phyloseq/tutorials-index) for processing Phyloseq objects can be found here. 

```{r}
## Install Phyloseq try http:// if https:// URLs are not supported
#source("https://bioconductor.org/biocLite.R")
# biocLite("phyloseq")

library(phyloseq)

biom_data <- import_biom(BIOMfilename = "Phyloseq/taxa_table.biom", treefilename = "Phyloseq/tree.nwk")

## The mapping file is still in the demo data as we did not operate on it
mapping_file <- import_qiime_sample_data(mapfilename = "Demo_Data/sample-metadata.tsv")

# Merge the OTU and mapping data into a phyloseq object
phylo <- merge_phyloseq(biom_data, mapping_file)
#Add names to biom table and check phyloseq objects
colnames(tax_table(phylo))= c("Kingdom","Phylum","Class","Order","Family","Genus", "Species")
rank_names(phylo)


depths <- sample_sums(phylo)
depths
```

And your all set! Good luck processing that data!


